### 2020-08-29 cases_and_policies/rmd/cf-pindex.Rmd
Added cumulative cases and deaths to counterfactuals.

---
title       : "Causal impact of masks, policies, behavior on early COVID-19 pandemic in the U.S."
subtitle    : "Counterfactuals (with policies averaged)"
author      :  Victor Chernozhukov,  Hiro Kasahara, Paul Schrimpf
job         :
date        : "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: "covid.bib"
link-citations: true
always_allow_html: true
output      :
    html_document :
        toc : true
        toc_depth : 3
        toc_float : true
        number_sections : true
        #theme : journal
        #css : 628notes.css
        code_folding: hide
        #lib_dir : deps
        self_cononontained : true
        fig_width: 8
        fig_height: 6
---

@@

# Counterfactuals

@@

```{r statesim, cache=FALSE}
infovars <- list(c("dlogdc", "logdc")) #,
#c("dlogdc", "logdc"))
#infovars2 <- list(c( "logdc"))
tvars <- "dlogtests"
xlist <-list("")
interactions <-list("month", statevars)
ilist <- list(interactions) #, interactions)
iv <- list("0") #,"(testratedc ~ lag(testratedc,14))")

L <- 14
sdf <- subset(df, df$date>=datelims[1]) #+L)
regs <- mainregressions2(sdf,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
m <- regs$piy[[1]]$reg
cnames <- c(pols, infovars[[1]])
di <- dieff_table2(lapply(regs$pib[[1]][1:4], function(x) x$reg$coef[,1]),
                  regs$pbiy[[1]]$reg$coef[,1],
                  regs$piy[[1]]$reg$coef[,1],
                  policies=cnames)

@@

simmodel <- createforwardsim(m, sdf)

# remove policies
cf <- sdf
cf[,pols] <- 0
nopmodel <- createforwardsim(m, cf)

rerunsim=FALSE
if (!file.exists("cfnop_idx.Rda") || rerunsim) {
  cfnoplist <- lapply(unique(sdf$state), function(s)
    nopstatesim(s,simobs=simmodel,simcf=nopmodel,data=sdf))
  allmeans <- rbindlist(lapply(cfnoplist, getmeans))
  save(cfnoplist,allmeans, file="cfnop_idx.Rda")
}
load("cfnop_idx.Rda")
cfnop <- do.call(rbind, lapply(cfnoplist, function(x) x$edf))
```

We now examine the impact changing from the observed policies to
none. For illustrative purpose, we being by focusing on
Washington. Figure \ref{fig:WA-nop-dlogdc} shows the observed,
estimated average, and counterfactual without policies average of
$\Delta \log \Delta C$ in Washington.

```{r WA-nop-dlogdc, cache=TRUE,fig.cap="Case growth without policies in Washington \\label{fig:WA-nop-dlogdc}"}
st <- "Washington"
f <- cfplots(st, cfnop, data=df)
grid.arrange(f$figd, f$figdp, nrow=2, top =
                                        textGrob(st,gp=gpar(fontsize=20,font=2)))

```{r WA-nop-cases, cache=TRUE, fig.cap="Cases without policies in Washington \\label{fig:WA-nop-cases}"}
grid.arrange(f$figl, f$figp, nrow=2) #,
#top = textGrob(st,gp=gpar(fontsize=20,font=2)))
```

```{r nop-all, cache=TRUE}
abbrv <- unique(df[,c("state","ST")])
cfnop <- merge(cfnop, abbrv, by="state")

figs <- nationalplots(allmeans,cfnop, "removing policies")

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_idx.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_idx.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_idx.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

@@

We now look at what happens when all states impose a mask mandate on
April 1st.

```{r mask_sim, cache=FALSE}
cf <- df
cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-04-01")
cf[,"pmaskbus"] <-
  smoothedpolicy(cf,"Mandate.face.mask.use.by.employees.in.public.facing.businesses",type="ma",bw=6)
cf <- subset(cf, cf$date>=datelims[1]) #+L)
maskmodel <- createforwardsim(m, cf)

@@

```{r, cache=TRUE}
cfmask <- merge(cfmask, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfmask, data=df)
  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on April 1st in %s",st)))

@@

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(maskall, cfmask, "mandating masks on April 1st\n")

@@

if (!file.exists("cfshelter_idx.Rda") || rerunsim) {
  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf, nresid=1))
  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
  save(cfshelterlist, shelterall, file="cfshelter_idx.Rda")
}
load("cfshelter_idx.Rda")
cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
```

@@

We now look at what happens when all states impose a mask mandate on
April 1st.

```{r mask_sim-deaths, cache=FALSE}
cf <- df
cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-04-01")
cf[,"pmaskbus"] <-
  smoothedpolicy(cf,"Mandate.face.mask.use.by.employees.in.public.facing.businesses",type="ma",bw=6)
cf <- subset(cf, cf$date>=datelims[1]) #+L)
maskmodel <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

@@

```{r, cache=TRUE}
cfmask <- merge(cfmask, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfmask, yvar="deaths", data=df)
  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on April 1st in %s",st)))

@@

figs <- nationalplots(maskall, cfmask, "mandating masks on April 1st\n", yvar="deaths")

@@

if (!file.exists("cfshelter_deaths_idx.Rda") || rerunsim) {
  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf, nresid=1))
  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
  save(cfshelterlist, shelterall, file="cfshelter_deaths_idx.Rda")
}
load("cfshelter_deaths_idx.Rda")
cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
```

>>>>

---
title       : "Causal impact of masks, policies, behavior on early COVID-19 pandemic in the U.S."
subtitle    : "Counterfactuals"
author      :  Victor Chernozhukov,  Hiro Kasahara, Paul Schrimpf
job         :
date        : "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: "covid.bib"
link-citations: true
always_allow_html: true
output      :
    html_document :
        toc : true
        toc_depth : 3
        toc_float : true
        number_sections : true
        #theme : journal
        #css : 628notes.css
        code_folding: hide
        #lib_dir : deps
        self_contained : true
        fig_width: 8
        fig_height: 6
---

@@

@@

```{r regressions, cache=FALSE}
infovars <- list(c("dlogdc", "logdc")) #,
#c("dlogdc", "logdc"))
#infovars2 <- list(c( "logdc"))
tvars <- "dlogtests"
xlist <-list("")
interactions <-list("month", statevars)
ilist <- list(interactions) #, interactions)
iv <- list("0") #,"(testratedc ~ lag(testratedc,14))")

L <- 14
sdf <- subset(df, df$date>=datelims[1]) #+L)
regs <- mainregressions2(sdf,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
m <- regs$piy[[1]]$reg
cnames <- c(pols, infovars[[1]])

di <- dieff_table2(lapply(regs$pib[[1]][1:4], function(x) x$reg$coef[,1]),
                  regs$pbiy[[1]]$reg$coef[,1],
                  regs$piy[[1]]$reg$coef[,1],
                  policies=cnames)

@@

rerunsim <- FALSE
```

# Method

We compute counterfactuals from the "total effect" version of the model, with behavior concentrated out.

$$
Y_{i,t+\ell}
= \mathsf{a}'
P_{it} + \mathsf{b}' I_{it} +  {\tilde {\delta}}' W_{it}   + {\bar \varepsilon}_{it}
$$

We consider a counterfactual change of $P_{it}$ to $P_{it}^*$, while
$W_{it}$ and $\bar \varepsilon_{it}$ are held constant. In response to
the policy change, $Y_{i,t+\ell}$ and the part of $I_{it}$ that
contains $Y_{it}$, change to $Y_{i,t+\ell}^*$ and $I_{it}^*$. To be
specific, let $C_{it}$ denote cumulative cases in state $i$ on day
$t$. Our outcome is:

$$
Y_{i,t+\ell} \equiv \log
\left(C_{i, t+\ell} - C_{i, t + \ell - 7} \right) - \log
\left(C_{i, t+\ell-7} - C_{i, t + \ell - 14} \right) \equiv \Delta
\log \Delta C_{i,t+\ell}
$$

where $\Delta$ is a 7 day difference operator. Writing the model in
terms of $\Delta \log \Delta C$ we have:

$$
\Delta \log \Delta C_{i,t+\ell} = \mathsf{a}' P_{it} + \mathsf{b}_D
\Delta \log \Delta C_{i,t} + \mathsf{b}_L \log \Delta C_{i,t} +
{\tilde {\delta}}' W_{it}   + {\bar \varepsilon}_{it}
$$

To simplify computation, we rewrite this model in terms of $\log
\Delta C$:

$$
\log \Delta C_{i,t+\ell} = \mathsf{a}' P_{it} + \log \Delta
C_{i,t+\ell-7} + (\mathsf{b}_D + \mathsf{b}_L) \log \Delta C_{i,t} -
\mathsf{b}_D \log \Delta C_{i,t-7} + {\tilde {\delta}}' W_{it}   + {\bar \varepsilon}_{it}
$$

This equation is used to iteratively compute
$\log \Delta C_{i,t+\ell}^*$
conditional on initial $\{\log \Delta C_{i,s}\}_{s=-\ell-7}^{0}$,
and the entire sequence of $\{W_{i,t}, P_{i,t}^*, \bar \varepsilon_{it} \}_{t=0}^T$.

Weekly cases instead of log weekly cases are given simply by

$$ \Delta C_{i,t+\ell}^* = \exp(\log \Delta C_{i,t+\ell}^*)$$

Cumulative cases can be recursively computed as:

$$
  C_{i,t} = C_{i,t-7} + \Delta C_{i,t} = C_{i,t-7} + \exp(\log \Delta C_{i,t})
$$

given initial $\{C_{i,s}\}_{s=-7}^0$.

Note that $\log \Delta C_{i,t+\ell}^*$ depends linearly on
$\bar \varepsilon$, so the residuals (and our decision to condition on
versus intergrate out) do not matter when considering linear
constrasts of $\log \Delta C_{it}$ or $\Delta \log \Delta C_{it}$, or
when considering relative constrasts of $\Delta C_{it}$.

On the other hand, cumulative cases are a nonlinear function of
$\bar\varepsilon$, so the conditioning on $\bar\varepsilon$ might
matter (not that there's anything wrong with that).

## Inference

Let $S(\theta, \mathbf{\varepsilon})$ denote some counterfactual
quantity or contrast of interest, where $\theta = (\mathsf{a},
\mathsf{b}, \tilde{\delta})$ are the parameters, and
$\mathbf{\varepsilon}$ is the vector of residuals. Examples of $S$
that we compute include:

- Contrasts of growth rates: $S(\theta, \mathbf{\varepsilon}) =
  \Delta \log \Delta C_{it}^* - \Delta \log \Delta C_{it}$

- Relative contrasts of weekly cases: $S(\theta, \mathbf{\varepsilon}) =
  \Delta C_{it}^*/\Delta C_{it}$

- Constrasts of cumulative cases: $S(\theta, \mathbf{\varepsilon}) =
  C_{it}^*/C_{it}$

The first two examples do not actually depend on
$\mathbf{\varepsilon}$, but the third one does. Inference is by
simulation. Let $\hat{\theta}$ denote our point estimates and
$\hat{\mathbf{\varepsilon}}$ the associated residuals. We draw
$\tilde{\theta}_j$ from the asymptotic distribution of
$\hat{\theta}$. Currently the code then computes

$$ \tilde{s}_j = S(\tilde{\theta}_j, \hat{\mathbf{\varepsilon}}) $$

for $j=1,..., 200$ and plots the mean across $j$ as a point estimate
and quantiles across $j$ for confidence intervals. This is fine when
$\varepsilon$ does not affect the quantity of interest, but I think
the correct thing to when $\varepsilon$ matters is to compute

$$ \tilde{s}_j = S(\tilde{\theta}_j, \tilde{\mathbf{\varepsilon}}_j) $$

where $\tilde{\mathbf{\varepsilon}}_j$ are the residuals from using
$\tilde{\theta}_j$ as coefficients.

# Cases

@@

We now look at what happens when all states impose a mask mandate on
May 14th.

```{r mask_sim, cache=FALSE}
sdf <- subset(df, df$date>=datelims[1]) #+L)
simmodel <- createforwardsim(m, sdf)
abbrv <- unique(df[,c("state","ST")])

cf <- df
cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-03-14")
cf[,"pmaskbus"] <-
  smoothedpolicy(cf,"Mandate.face.mask.use.by.employees.in.public.facing.businesses",type="ma",bw=6)
cf <- subset(cf, cf$date>=datelims[1]) #+L)
maskmodel <- createforwardsim(m, cf)

@@

```{r, cache=FALSE}
cfmask <- merge(cfmask, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfmask, data=df)
  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on May 14th in %s",st)))

@@

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_idx.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()

  if (st=="Washington") for (fig in f) print(fig)
}

figs <- nationalplots(maskall, cfmask, "mandating masks on May 14th\n")

@@

if (!file.exists("cfshelter_idx.Rda") || rerunsim) {
  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
  save(cfshelterlist, shelterall, file="cfshelter_idx.Rda")
}
load("cfshelter_idx.Rda")
cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
```

@@

We now look at what happens when all states impose a mask mandate on
May 14th.

```{r mask_sim-deaths, cache=FALSE}
cf <- df
cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-03-14")
cf[,"pmaskbus"] <-
  smoothedpolicy(cf,"Mandate.face.mask.use.by.employees.in.public.facing.businesses",type="ma",bw=6)
cf <- subset(cf, cf$date>=datelims[1]) #+L)
maskmodel <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

@@

```{r, cache=TRUE}
cfmask <- merge(cfmask, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfmask, yvar="deaths", data=df)
  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on May 14th in %s",st)))

@@

figs <- nationalplots(maskall, cfmask, "mandating masks on May 14th\n", yvar="deaths")

@@

if (!file.exists("cfshelter_deaths_idx.Rda") || rerunsim) {
  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
  save(cfshelterlist, shelterall, file="cfshelter_deaths_idx.Rda")
}
load("cfshelter_deaths_idx.Rda")
cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
```
